Operator evolution from the similarity renormalization group and the Magnus expansion

### 2020-06-19 Notes/Operator_evolution/operator_evolution.tex
Adding Overleaf files to repository for operator evolution project
This has all the figures and Latex files for "Operator evolution from the SRG and the Magnus expansion".

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% File: operator_evolution.tex
%
% Author: A. J. Tropiano (tropiano.4@osu.edu)
% Date: August 23, 2019
%
% Draft of paper on SRG operator evolution and the Magnus expansion.
%
% Revision history:
%	08/28/19 --- Started outlining and added pieces from old notes.
%	09/27/19 --- Wrote the introduction and swapped the Magnus and operator sections.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\documentclass[preprintnumbers,floatfix,aps,prc,preprint,nofootinbib]{revtex4-1}

% Packages
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{bm}
\usepackage[font=small,skip=0pt]{caption} % For captions on figures and tables
\usepackage{cellspace}
\usepackage{color}
\usepackage{enumerate}
\usepackage{epsfig}
\usepackage[figuresright]{rotating}
\usepackage{float}
\usepackage{hyperref} % For clickable links to sections within table of contents
\usepackage{graphicx}
\graphicspath{{../../Figures/}} % Setting the graphics path
\usepackage{physics} % For bra-ket notation
\usepackage{siunitx}
\usepackage[caption=false]{subfig} % For sub-figures

\newcommand{\eps}{\varepsilon}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\title{Operator evolution from the similarity renormalization group and the Magnus expansion}

@@

\begin{abstract}

\noindent{Ideas for Magnus / SRG operator evolution paper}
\\
-- SRG/Magnus evolution in different potentials (non-local, local, semi-local). Universality. High cutoffs.
\\
-- Block-diagonal generator for high cutoff potentials and operator evolution. How the block-diagonal generator handles spurious bound states.
\\
-- Testing the Magnus expansion for high cutoff potentials using the potentials from Wendt 2011 for comparison. Spurious bound states and connection to intruder states in IMSRG calculations.
\\
-- Operator evolution for different potentials and generators.

\end{abstract}

@@

% Unnecessary paragraph.
%\noindent{%
%\textbf{NN interaction and SRG decoupling.}
%}
%\\
%The nuclear many-body problem has long been hindered by the short-range repulsive core and strong tensor force in nucleon-nucleon (NN) interactions leading to non-perturbative few- and many-body systems and strongly correlated wave functions. The Fourier transform of NN interactions gives rise to the coupling of low- and high-momentum modes, which is manifested as non-zero off-diagonal potential matrix elements. It is very difficult to implement these interactions in many-body methods using basis expansions, because the matrix dimension becomes too large for accurate calculations. Renormalization group (RG) methods can be used to soften the interaction to make many-body methods feasible. One such method, the similarity renormalization group (SRG), allows for a continuous change in resolution of the potential without changing the observables. More generally, the SRG can evolve any operator and wave function for consistent calculation of observables.
%\\

Nuclear physics has seen a massive increase in the number of new NN interactions from chiral effective field theory ($\chi^{EFT}$) in the last decade. Previous SRG studies were limited to phenomenological interactions or some of the first non-local chiral interactions \cite{Entem:2003ft}, but with much more variety in modern chiral interactions, we would like to revisit the question of how different potentials and other operators change under SRG transformations. Since $\chi^{EFT}$ is an effective field theory, different potentials will still give the same $S$-matrix, that is, the potentials are phase equivalent up to some higher energy. However, the potentials themselves can differ in many ways such as the regulator functions, cutoff, order, and so forth, and thus, the potential matrix elements are not the same. Dainton \textit{et al.}~\cite{Dainton:2013axa} demonstrated SRG transformations drive different NN potentials to the same low-momentum matrix elements, that is, a flow to universality up to the momentum value of phase inequivalence. We address the question of universality for modern chiral potentials as well as how other operators change with respect to the corresponding SRG transformations.
\\

The SRG decouples low- and high-momentum scales in the Hamiltonian by applying a continuous unitary transformation $U(s)$ where $s=0 \rightarrow \infty$ is the flow parameter. An evolved operator is given by
%
\begin{eqnarray}
	\label{eq:srg_operator}
	O(s) = U(s) O(0) U^{\dagger}(s),
\end{eqnarray}
%
where $O(0)$ corresponds to the initial operator. Because $U(s)$ is unitary, the observables of the operator are preserved. In practice, the unitary transformation U(s) is not explicitly solved for; the evolved operator is given by solving a differential flow equation which is obtained by taking the derivative of Eqn. (\ref{eq:srg_operator}),
%
\begin{eqnarray}
	\label{eq:srg_flow}
	\frac{dO(s)}{ds} = [\eta(s), O(s)],
\end{eqnarray}
%
where $\eta(s)=\frac{dU(s)}{ds} U^{\dagger}(s) = -\eta^{\dagger}(s)$ is the anti-hermitian SRG generator. The generator is defined as a commutator, $\eta(s) = [G, H(s)]$, where $G$ specifies the type of flow and $H(s)$ is the evolved Hamiltonian. The choice in $G$ determines the pattern of decoupling in the Hamiltonian.
\\

By setting $G=H_D(s)$, the diagonal of the Hamiltonian, the Hamiltonian is driven to band-diagonal form. This choice was implemented by Wegner in condensed matter physics \cite{Wegner:1994ab}. In a similar option used in nuclear physics, $G$ is set to the relative kinetic energy, $T_{rel}$, which also drives to band-diagonal form. In most cases, these two choices give the same evolved operators. In exceptional cases, which will be discussed later in the paper, the two band-diagonal choices can have drastically different behavior. For band-diagonal decoupling, it is convenient to define $\lambda \equiv s^{-1/4}$ which roughly measures the width of the band-diagonal in the decoupled operator. For block-diagonal decoupling \cite{Anderson:2008mu}, 
%
\begin{eqnarray}
	\label{eq:block_diag_g}
	G =
	\begin{pmatrix}
		PH(s)P & 0 \\
		0 & QH(s)Q \\
	\end{pmatrix}
	\equiv H_{BD}(s),
\end{eqnarray}
%
where $P$ and $Q$ are low- and high-momentum projection operators, the Hamiltonian is split into low- and high-momentum sub-blocks by specifying a separation in momentum $\Lambda$ and taking $s \rightarrow \infty$. In momentum-space, the projection operators are step functions defined by the sharp cutoff in $\Lambda$. These transformations are similar to V\textsubscript{low $k$} transformations \cite{Bogner:2001gq, Bogner:2001jn, Bogner:2003wn} but keep the high-momentum matrix elements non-zero, although entirely decoupled from the low-momentum sub-block. Generally, the flow equation (\ref{eq:srg_flow}) is solved up to some finite value of $s$ with a high-order ODE solver. For notational convenience, we write the generators without the $s$ dependence in the rest of the paper.
\\

The decoupling of matrix elements does not necessarily hold in other operators as it does in the Hamiltonian. In a previous work \cite{Anderson:2010aq}, it was shown that SRG transformations induce low-momentum contributions in high-momentum operators and change very little in low-momentum operators. We investigate whether this is a general trend of the SRG for any potential and SRG generator. Instead of solving another set of flow equations, we apply SRG band- and block-diagonal transformations to other operators by building the unitary transformations explicitly using the eigenvectors of the initial and evolved Hamiltonians. SRG-evolved operators depend on the NN interaction despite having the same initial operator, since the flow equation (\ref{eq:srg_flow}) depends on $H(s)$ from the definition of $\eta(s)$.
\\

There has also been renewed interest in studying chiral interactions at high cutoffs \cite{Tews:2018sbi}. In spin-triplet channels and at leading-order (LO), taking the cutoff to high values can result in spurious, deeply bound states from a highly singular short-ranged tensor force. In principle, these deep bound states are outside the range of the EFT and are absorbed by counterterms at higher order. In \cite{Wendt:2011qj}, it was shown that band-diagonal SRG decoupling of NN potentials in partial waves with spurious bound states differs significantly depending on the choice of SRG generator. These high cutoff potentials provide a unique opportunity to understand some of the previous points about universality, the choice of SRG generator and decoupling, as well as a good test problem for a variant of the SRG, which utilizes the Magnus expansion \cite{Morris:2015yna}.
\\

The Magnus expansion gives the SRG the capability to solve for the unitary transformation explicitly which can then be applied to any other operator of interest. This offers a huge numerical simplification in Fock-space where the matrix size of operators can become very large. The Magnus expansion is often used in in-medium SRG (IMSRG) calculations to solve the nuclear many-body problem directly. In some IMSRG calculations, intruder states can shift to the energy scale of the reference state in valence-space and severely distort low-energy properties. It is not understood how the IMSRG procedure evolves intruder state systems. Due to the prominence of the Magnus expansion in IMSRG, one must first understand if the Magnus approach matches the typical approach for difficult systems before investigating the intruder state problem. For this reason, NN potentials with spurious bound states offer a good test laboratory for the Magnus implementation to better understand if the Magnus is equivalent to the SRG approach.
\\

\noindent{%
\textbf{Overview of sections.}
}

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% File: operator_evolution.tex
%
% Authors: Anthony Tropiano (tropiano.4@osu.edu),
%          Dick Furnstahl (furnstahl.1@osu.edu),
%          Scott Bogner (bogner@frib.msu.edu).
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\documentclass[10pt,aps,prc,floatfix,twocolumn,nofootinbib]{revtex4-1}
%\documentclass[preprintnumbers,floatfix,aps,prc,preprint,nofootinbib]{revtex4-1}

% Packages
\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{bbold}
\usepackage{bm}
\usepackage{cellspace}
\usepackage{color}
\usepackage{enumerate}
\usepackage{graphicx}
\graphicspath{{Figures/}} % Setting the graphics path
\usepackage{hyperref} % For clickable links to sections within table of contents
\usepackage{physics} % For bra-ket notation
\usepackage[figuresright]{rotating}
\usepackage{siunitx}
\usepackage[caption=false]{subfig} % For sub-figures
%\usepackage{floatrow}

% Macros
\newcommand{\chiEFT}{\ensuremath{\chi{\textrm{EFT}}}}
\newcommand{\LambdaBD}{{\Lambda_{\text{BD}}}}
\newcommand{\Trel}{\ensuremath{T_{\textrm{rel}}}}
\newcommand{\ataq}{a^{\dagger}_q a_q}
\newcommand{\abi}{\textit{ab initio}}

\newcommand{\mystrut}{\rule{0pt}{0.9\normalbaselineskip}}
\newcommand{\Klo}{K_{\text{lo}}}
\newcommand{\Khi}{K_{\text{hi}}}
\newcommand{\dst}{\displaystyle}

\newcommand{\PP}{\ensuremath{P\mbox{--}P}}
\newcommand{\PQ}{\ensuremath{P\mbox{--}Q}}
\newcommand{\QP}{\ensuremath{Q\mbox{--}P}}
\newcommand{\QQ}{\ensuremath{Q\mbox{--}Q}}

%\floatsetup[figure]{style=plain, subcapbesideposition=top}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\title{Operator evolution from the similarity renormalization group \texorpdfstring{\\}{}
and the Magnus expansion}

@@

\begin{abstract}
The Magnus expansion is an efficient alternative to solving similarity renormalization group (SRG) flow equations with high-order, memory-intensive ordinary differential equation solvers. The numerical simplifications it offers for operator evolution are particularly valuable for in-medium SRG calculations, though challenges remain for difficult problems involving intruder states. Here we test the Magnus approach in an analogous but more accessible situation, which is the free-space SRG treatment of the spurious bound-states arising from a leading-order chiral effective field theory (EFT) potential with very high cutoffs. We show that the Magnus expansion passes these tests and then use the investigations as a springboard to address various aspects of operator evolution that have renewed relevance in the context of the scale and scheme dependence of nuclear processes. These aspects include SRG operator flow with band- versus block-diagonal generators, universality for chiral EFT Hamiltonians and associated operators with different regularization schemes, and the impact of factorization arising from scale separation. Implications for short-range correlations physics and the possibilities for reconciling high- and low-resolution treatments of nuclear structure and reactions are discussed.
\end{abstract}

@@

Similarity renormalization group (SRG) transformations are a valuable tool for low-energy nuclear physics, whether applied in free space to soften input Hamiltonians for few- and many-body calculations, or for in-medium SRG (IMSRG) calculations that directly target the ground state or low-lying states in a given nucleus ~\cite{Bogner:2009bt,Furnstahl:2013oba,Hergert:2016iju}.
For both free-space and in-medium formulations, it is imperative that other operators are consistently and accurately evolved so that measurable quantities are left invariant.  
In the present work, we address the robustness of the Magnus expansion as a method to solve free-space SRG flow equations, and examine other issues of SRG operator evolution in light of the proliferation of new chiral EFT (\chiEFT) interactions~\cite{Epelbaum:2014efa,Gezerlis:2014zia,Piarulli:2014bda,Ekstrom:2015rta,Carlsson:2015vda,Reinert:2017usi,Ekstrom:2017koy,Entem:2017gor}, the scale dependence of short-range-correlation (SRC) physics~\cite{Hen:2016kwk,Weiss:2016obx,Cruz-Torres:2019fum,Schmidt:2020kcl}, recent interest in high-cutoff effective field theories (EFTs) and renormalization~\cite{Tews:2018sbi,Hammer:2019poc,vanKolck:2020llt,Tews:2020hgp}, and the universality of evolved operators~\cite{Dainton:2013axa,Arriola:2016fkr}.

The SRG decouples low- and high-momentum scales in a Hamiltonian by applying a continuous unitary transformation $U(s)$, where $s=0 \rightarrow \infty$ is the flow parameter~\cite{Bogner:2006pc}.
An evolved operator is given by
%
\begin{eqnarray}
	\label{eq:srg_operator}
	O(s) = U(s) O(0) U^{\dagger}(s),
\end{eqnarray}
%
where $O(0)$ is the initial operator.
Because $U(s)$ is unitary, matrix elements of the operator in evolved states are preserved.
An evolved operator can be found by solving a differential flow equation obtained by taking the derivative of Eq.~\eqref{eq:srg_operator},
%
\begin{eqnarray}
	\label{eq:srg_flow}
	\frac{dO(s)}{ds} = [\eta(s), O(s)],
\end{eqnarray}
%
where $\eta(s)=\frac{dU(s)}{ds} U^{\dagger}(s) = -\eta^{\dagger}(s)$ is 
the anti-hermitian SRG generator.
For the free-space SRG, the generator is typically defined as a commutator, $\eta(s) = [G, H(s)]$, where $G$ specifies the type of flow.
The choice of $G$ determines the pattern of decoupling in the Hamiltonian.

By setting $G=H_D(s)$, the diagonal of the Hamiltonian, the Hamiltonian is driven to band-diagonal form~\cite{Wegner:1994ab}.
In low-energy nuclear physics, $G$ is usually taken to be the relative kinetic energy, \Trel; i.e., the diagonal of the potential is not included in $G$. 
In most nuclear physics applications these two choices give the same evolved operators. 
But in exceptional cases involving evolution across bound states, which we consider in the next section, the two band-diagonal choices can have drastically different behaviors~\cite{Glazek:2008pg,Wendt:2011qj}. 
For band-diagonal decoupling, it is convenient to define $\lambda \equiv s^{-1/4}$, which roughly measures the width of the band-diagonal in the decoupled Hamiltonian~\cite{Anderson:2010aq}.

For block-diagonal decoupling~\cite{Anderson:2008mu,Szpigel:2016rbt}, $G$ is formed by splitting the Hamiltonian into low- and high-momentum sub-blocks as specified by a momentum separation scale $\LambdaBD$,
%
\begin{eqnarray}
	\label{eq:block_diag_g}
	G =
	\begin{pmatrix}
		PH(s)P & 0 \\
		0 & QH(s)Q \\
	\end{pmatrix}
	\equiv H_{BD}(s).
\end{eqnarray}
%
Here $P$ and $Q$ are low- and high-momentum projection operators.
In momentum space, the projection operators are step functions defined by the sharp cutoff $\LambdaBD$, although smoothed versions are also possible and may be preferred in some applications to avoid numerical artifacts~\cite{Anderson:2008mu}. 
These transformations are similar to V\textsubscript{low $k$} transformations~\cite{Bogner:2001gq,Bogner:2001jn,Bogner:2003wn} but keep the high-momentum matrix elements non-zero, maintaining a unitary transformation in the full space.
Complete decoupling of the blocks is in principle only reached in the $s \rightarrow \infty$ limit. 
In practice it is sufficient to solve the flow equation \eqref{eq:srg_flow} up to some finite value of $s$ with a high-order ODE solver such that the remaining ``neck'' between blocks is much narrower than $\LambdaBD$.

The SRG procedure can be implemented by solving the flow equation Eq.~\eqref{eq:srg_flow} for the evolved Hamiltonian simultaneously with other operators of interest.
However, one can also solve Eq.~\eqref{eq:srg_flow} exclusively for the evolved Hamiltonian and build the unitary transformation directly using the eigenvectors of the evolved and initial Hamiltonians (as is done in Sec.~\ref{sec:evolution_other_operators}).
Another approach is to solve the following equation for the unitary transformation
%
\begin{eqnarray}
	\label{eq:unitary_trans}
	\frac{dU(s)}{ds} = \eta(s) U(s),
\end{eqnarray}
%
which arises in an intermediate step in deriving Eq.~\eqref{eq:srg_flow}.
This is the starting point in the Magnus expansion implementation of the SRG.

The Magnus expansion gives us the capability to solve for the SRG unitary transformation with negligible violations of unitarity from numerically solving the ODEs,%
\footnote{There is a small numerical violation of unitarity in the standard approach to solving SRG equations due to accumulated time-step errors. With the Magnus expansion, unitarity is preserved to much higher precision because of the form of the transformation, as detailed in Sec.~\ref{sec:magnus_expansion_formalism}.}
after which it can be applied to any other operator of interest~\cite{Morris:2015yna}.
By utilizing an exponential parameterization for the transformation, $U(s)=e^{\Omega(s)}$, Eq.~\eqref{eq:unitary_trans} is recast as a flow equation for the anti-Hermitian operator $\Omega(s)$.
The solution of the flow equation for $\Omega(s)$ permits the use of cheap low-order ODE methods since the exponentiated operator is still unitary even if it has accumulated non-negligible time-step errors~\cite{Morris:2015yna}.
The Magnus expansion also offers important advantages over the direct solution of Eq.~\eqref{eq:unitary_trans} in Fock space, where practical calculations require operators to be truncated at the $a-$body level ($a<A$).
For instance, even if the Magnus flow equations are truncated at the two-body level, the resulting unitary transformation contains higher-body components from the exponentiation of $\Omega$.%
\footnote{This is similar to the advantages of truncated Coupled Cluster theory calculations relative to truncated Configuration Interaction calculations.}

Due to these advantages, most large-scale IMSRG calculations now utilize the Magnus expansion.
There are still open problems though.
For instance, in applications of the IMSRG to derive effective valence shell model Hamiltonians in multi-shell valence spaces, intruder states, which are low-lying states whose wave functions are dominated by high-energy configurations outside the model space, can severely distort low-energy properties or even prevent the flow from converging.
It is not yet fully understood how the IMSRG procedure evolves intruder state systems, though it appears that induced three- and higher-body operators rapidly grow in size for such systems, destroying the cluster hierarchy ($2N \gg 3N \gg 4N \gg \ldots$) in the evolved Hamiltonian~\cite{Stroberg:2019mxo}.

Interestingly, there is an analog to the intruder state problem in the much simpler two-nucleon problem.
In spin-triplet channels and at leading-order (LO) in \chiEFT, taking the EFT cutoff to high values can result in spurious, deep-bound states due to the highly singular short-ranged tensor force from one-pion exchange. 
In principle, these deep-bound states are not a problem because they are outside the range of the EFT.
In practice, there are subtleties analogous to the intruder state problem when one attempts to soften such Hamiltonians with free space SRG evolution.
In Ref.~\cite{Wendt:2011qj}, it was shown that band-diagonal SRG decoupling of NN potentials in partial waves with spurious bound states fails for the standard $G=\Trel$ generator, as the flow forces the deep bound state into the low-momentum sector.
As a result, there is no decoupling of high- and low-momentum physics, and the evolved interactions become increasingly singular at low momentum.
In contrast, the Wegner generator $G=H_{D}$ succeeds at depositing the spurious state(s) along the diagonal in the high-momentum sector, which is more natural as it allows a clean decoupling of high- and low-momentum physics. 
Since these findings were for the direct solution of Eq.~\eqref{eq:srg_flow}, this  provides a good test case for the Magnus approach and we document its performance in detail.
More generally, there has been renewed interest in studying chiral interactions at high cutoffs~\cite{Tews:2018sbi}. 
These high-cutoff chiral potentials provide us a laboratory to explore the effects of the SRG generator on decoupling, universality, and SRCs.
These issues also inform the behavior of standard \chiEFT\ potentials.

While interactions from \chiEFT\ have become the standard choice for \abi\ calculations of nuclei, they are not unique, even when restricted to the commonly used Weinberg power counting, because of many choices for regularization schemes and fitting protocols and even degrees of freedom (i.e., with or without Deltas).
In the early applications of \chiEFT\ potentials to nuclei, these choices were not explored but in recent years there has been a proliferation of nucleon-nucleon (NN) potentials and associated three-nucleon forces (e.g., see Refs.~\cite{Epelbaum:2014efa,Gezerlis:2014zia,Piarulli:2014bda,Ekstrom:2015rta,Carlsson:2015vda,Reinert:2017usi,Ekstrom:2017koy,Entem:2017gor}).
This diversification motivates us to revisit SRG operator evolution.
Past studies were limited to phenomenological interactions or a single class of chiral interactions (namely the non-local-regulated potentials from Refs.~\cite{Entem:2003ft} or \cite{Epelbaum:2004fk}).
Here we examine the fate of scheme dependence for new-generation NN potentials and associated operators as they are evolved to lower resolutions.

One intriguing aspect is universality.
By virtue of fitting to the same data or phase shifts, different \chiEFT\ potentials generate close to the same $S$-matrix in the energy range where there is a good fit; that is, the potentials are phase equivalent in that range.
However, matrix elements of the potentials in momentum space differ significantly based on the EFT order and the choice of regulator function and cutoff (scale and scheme dependence). 
Nevertheless, it has been observed that SRG transformations drive different NN potentials toward the same low-momentum matrix elements; in particular, this flow to universality is seen up to the momentum value of phase inequivalence~\cite{Bogner:2003wn,Bogner:2001gq,Dainton:2013axa}. 
We examine whether universality holds for modern chiral potentials but also address universality for other operators evolving under the corresponding SRG transformations.
This has implications for the analysis of reactions at different resolution scales~\cite{Furnstahl:2013dsa,More:2015tpa,More:2017syr}.

The question of whether non-Hamiltonian operators decouple or take universal forms has not been fully addressed in the literature.%
\footnote{Note that if the wave functions are decoupled, it is not necessary for the operators themselves to decouple to get decoupled matrix elements. See examples below.}
In fact, the decoupling of matrix elements does not necessarily result for other operators as it does for the Hamiltonian.
In previous work \cite{Anderson:2010aq,Bogner:2012zm}, it was found that SRG evolution induces low-momentum contributions in high-momentum operators and changes low-momentum operators very little, as might be expected from general EFT considerations.
We investigate whether this is a general trend of the SRG for a wider selection of potentials and SRG generators, explicitly analyze the nature of the evolution for representative high-momentum and low-momentum operators, and relate these observations to the high-resolution picture of SRCs and the role of factorization.

The plan of the paper is as follows.
We first revisit the high-cutoff problem, and test the Magnus approach in Sec.~\ref{sec:high_cutoffs_magnus_exp}.
We consider evolution of new-generation NN Hamiltonians in Sec.~\ref{sec:srg_evolution_nn_potentials} and then turn to other operators in Sec.~\ref{sec:evolution_other_operators}.
Our conclusions and outlook are summarized in Sec.~\ref{sec:summary}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{High cutoffs and the Magnus expansion}
\label{sec:high_cutoffs_magnus_exp}

%........................................................................
\subsection{High cutoffs and spurious bound states}
\label{sec:high_cutoffs}

%
\begin{figure*}[tbh]
    \includegraphics[clip,width=0.8\textwidth]{potential_slices_3S1_kvnns_900_901_902.pdf}
	\caption{Diagonal and far off-diagonal matrix elements of non-local LO potentials at cutoffs $\Lambda=4$ (black), $9$ (red) and $20$\,fm$^{-1}$ (blue), SRG-evolving left to right under transformations with Wegner (solid) and block-diagonal (dashed) generators in the $^3$S$_1$ channel.
	We vary the SRG flow parameter $\lambda$ for Wegner evolution and fix it at $\lambda=1.2$\,fm$^{-1}$ for block-diagonal evolution. The decoupling scale in the block-diagonal generator is denoted by $\LambdaBD$.}
	\label{fig:potential_slices_high_cutoffs}
\end{figure*}
%

The \chiEFT\ potentials used in most \abi\ nuclear calculations are not renormalizable in the sense that dependence on the regulator is not suppressed by taking the momentum cutoff increasingly high (or low, if a coordinate-space regulator).
However, Nogga \textit{et al.}~\cite{Nogga:2005hy} showed that the LO version of these interactions, with promoted counterterms in some channels, \emph{is} renormalizable in this sense.
There is active work on renormalizable power counting for \chiEFT\ beyond LO (see references cited in \cite{Tews:2018sbi}).

The LO theory at high cutoff is a useful laboratory for testing the SRG (as well as providing insight into the evolution of SRC physics, see Sec.~\ref{sec:evolution_other_operators}).
It features the appearance of spurious, deeply bound states in some channels, which is ultraviolet physics beyond the range of the EFT, and thus does not violate EFT principles.
However, these present a major challenge to the SRG.
Wendt \textit{et al.} studied SRG band-diagonal transformations of high-cutoff LO potentials~\cite{Wendt:2011qj} and showed that channels with spurious deep-bound states did not automatically exhibit the expected decoupling and universality of the potential if the conventional SRG generator is used.
In particular, the spurious bound state(s) is driven from high to low momentum in the evolved potential when applying transformations with $G=\Trel$.
The observables remain unchanged because the transformation is still unitary, but the potential and wave functions are altered significantly by the presence of the spurious bound state at low momentum. 
In contrast, if the Wegner generator is used, the spurious state(s) is decoupled, subsequently yielding universality in low-momentum matrix elements of the potential.

In Fig.~\ref{fig:potential_slices_high_cutoffs} we show SRG band- and block-diagonal evolution of high-cutoff non-local potentials at LO, 
which consists of one-pion exchange and a contact interaction.
We restrict our attention to the $^3$S$_1$--$^3$S$_1$ sub-block of the coupled $^3$S$_1$--$^3$D$_1$ channel (note that spurious, deeply bound states only appear in spin-triplet channels~\cite{Nogga:2005hy}).
The contact interaction is determined by fitting the associated low-energy constant to $E_{\text{lab}}=10$ MeV phase shift data.
In the diagonal matrix elements of Fig.~\ref{fig:potential_slices_high_cutoffs}, we see a steep drop-off in the Wegner transformed potentials around $k \approx 1.6-1.75$\,fm$^{-1}$.
This corresponds to the decoupled spurious bound state ($\varepsilon \approx -2000$ MeV).
The value of momentum where the spurious bound state decouples is a scheme-dependent quantity that is sensitive to how momentum space is discretized (the momentum mesh).
Due to this dependence, we have been unable to predict the value of $k$ at which the spurious state decouples.
However, when the spurious state is decoupled outside the low-momentum part of the potential, the Wegner evolution collapses the low-momentum matrix elements of the different potentials to the same mesh-independent values in accordance with universality.
There is no drop-off in the $\Lambda=4$\,fm$^{-1}$ potential as it has no spurious state.

We also see universality for the block-diagonal evolved matrix elements as before, but there is no noticeable influence from the spurious bound state.
This is due to the band-diagonal generator locally decoupling the matrix elements whereas the block-diagonal generator cleanly separates the potential into a low-momentum sub-block and a high-momentum sub-block.
In the limit $\lambda \rightarrow 0$ with $\LambdaBD$ sufficiently low, the spurious deep-bound state(s) is contained entirely in the high-momentum sub-block.
We verified this by diagonalizing the sub-blocks separately to see which one contained the spurious bound state.
We identify $\LambdaBD \approx 4.5$\,fm$^{-1}$ as the approximate value at which the spurious bound state switches from the low-momentum sub-block to the high-momentum sub-block.
Thus, the block-diagonal transformations decouple the spurious state(s) at a higher value of momentum than the Wegner transformations, which isolates the physical states more effectively.
Tests with different meshes found the same value of momentum, suggesting a scheme-independent result.
However, we do not have an analytic understanding of these scales.

We can draw a loose analogy to intruder states corrupting low-energy physics in IMSRG calculations with spurious bound states corrupting universality in SRG-evolved potentials.
From the analysis so far, it is evident that the choice of SRG generator is important in properly decoupling the high-momentum spurious state from low-momentum physics.
It would be interesting to generalize this conclusion to an $A$-body system and analyze how different generators deal with intruder states.
However, we must first verify that the Magnus approach is the same as the conventional SRG approach for difficult systems because of its use in IMSRG calculations.
In the following sub-sections, we present the Magnus approach and compare to the conventional SRG using high-cutoff potentials with spurious states as a test case.

%........................................................................
\subsection{The Magnus expansion: Formalism}
\label{sec:magnus_expansion_formalism}

We briefly review the formalism of the Magnus expansion and its use in the SRG.
Mathematically speaking, the Magnus expansion is a method for solving an initial value problem associated with a linear ordinary differential equation.
Formal details of the Magnus expansion are discussed in \cite{Blanes:2009ab}.
We will introduce the Magnus expansion in the context of SRG operator evolution.

%
\begin{figure*}[tbh]
	\includegraphics[clip,width=0.8\textwidth]{potential_slices_kvnn_901_Wegner_k_magnus_values_2_6_10.pdf}%
	\caption{Comparison of SRG- and Magnus-evolved diagonal and far off-diagonal matrix elements of the non-local LO potential (9\,fm$^{-1}$) in the $^3$S$_1$ channel for several truncations $k_{\rm max}$ in the Magnus sum \eqref{eq:magnus_omega}. Here we evolve in $\lambda$ left to right using the Wegner generator $G=H_D$.}
	\label{fig:potential_slices_high_cutoffs_Wegner}
\end{figure*}
%
\begin{figure*}[tbh]
	\includegraphics[clip,width=0.8\textwidth]{potential_slices_kvnn_901_T_k_magnus_values_2_6_10.pdf}%
	\caption{Same as in Fig.~\ref{fig:potential_slices_high_cutoffs_Wegner} but with $G=\Trel$.}
	\label{fig:potential_slices_high_cutoffs_T}
\end{figure*}
%

We can solve Eq.~\eqref{eq:unitary_trans} with a solution $U(s)=e^{\Omega(s)}$ where $\Omega^{\dagger}(s) = -\Omega(s)$, and $\Omega(0) = 0$.
$\Omega(s)$ is expanded as a power series in $\eta(s)$:
%
\begin{eqnarray}
    \label{eq:magnus_expansion}
    \Omega(s) = \sum_{n=1}^{\infty} \Omega_n(s),
\end{eqnarray}
%
where the terms of the series are given by integral expressions involving $\eta(s)$,
%
\begin{align}
    \label{eq:omega_terms}
    \Omega_1(s) &= \int_0^s\, ds_1 \eta(s_1), \nonumber \\        
    \Omega_2(s) &= \frac{1}{2} \int_0^s ds_1 \int_0^{s_1} ds_2 \, [\eta(s_1), \eta(s_2)], \\
    & \quad\vdots \nonumber
\end{align}
%
Equation~\eqref{eq:magnus_expansion} is referred to as the Magnus expansion.
(Again, see \cite{Blanes:2009ab, Magnus:1954zz} for further
details.)
We avoid computing the integral terms $\Omega_n(s)$ since it requires storing $\eta(s)$ over a range of $s$ values, which is impractical for large-scale calculations.
We focus instead on the derivative of $\Omega(s)$,
%
\begin{eqnarray}
	\label{eq:magnus_omega}
	\frac{d\Omega(s)}{ds} = \sum_{k=0}^{\infty} \frac{B_k}{k!} ad_{\Omega}^{k}(\eta),
\end{eqnarray}
%
where $B_k$ are the Bernoulli numbers, $ad_{\Omega}^{0}(\eta)=\eta(s)$, and $ad_{\Omega}^{k}(\eta)=[\Omega(s),ad_{\Omega}^{k-1}(\eta)]$.
We integrate this differential equation to find $\Omega(s)$ and evaluate the unitary transformation.
Then the evolved operator can be evaluated with the Baker-Cambell-Hausdorff formula~\cite{Morris:2015yna},
%
\begin{eqnarray}
	\label{eq:bch}
	O(s) = e^{\Omega(s)} O(0) e^{-\Omega(s)} = \sum_{k=0}^{\infty} \frac{1}{k!} ad_{\Omega}^{k}(O).
\end{eqnarray}
%
As $k \rightarrow \infty$ in both sums in Eqs.~\eqref{eq:magnus_omega} and \eqref{eq:bch}, the Magnus transformation matches the SRG transformation exactly.%
\footnote{Note that this equivalence is exact only if both series converge and the ODEs in Eq.~\eqref{eq:magnus_omega} are solved exactly.}
We investigate several truncations $k_{\rm max}$ in Eq.~\eqref{eq:magnus_omega} and take many terms, $k_{\rm max} \sim 25$, in Eq.~\eqref{eq:bch}.

%
\begin{figure*}[tbh]
	\includegraphics[clip,width=0.67\textwidth]{eta_omega_norms_kvnns_900_901_902_3S1_Wegner_k_magnus_6_ds1,0e-05.pdf}%
	\caption{Frobenius norms of $\eta(s)$ and $\Omega(s)$ from Magnus-evolving the high-cutoff LO potentials in the $^3$S$_1$--$^3$D$_1$ coupled channel: $\Lambda=4$ (black dashed), $9$ (red dash-dotted) and $20$\,fm$^{-1}$ (blue solid) where $G=H_D$ and $k_{\rm max}=6$.}
	\label{fig:eta_omega_norms}
\end{figure*}
%

There are significant advantages to the Magnus implementation in IMSRG calculations.
In the conventional approach, the numerical error associated with solving Eq.~\eqref{eq:srg_flow} accumulates directly in the operator and can distort the eigenvalues of the transformed Hamiltonian.
To guard against this, one must use a high-order ODE solver, which can become prohibitive for large-scale calculations due to the memory-intensive nature of such solvers.  
In the Magnus implementation, unitarity is guaranteed by the form of $U(s)$.
One can solve Eq.~\eqref{eq:magnus_omega} with a low-order stepping method with a substantially lower memory footprint, which nevertheless preserves the eigenvalues exactly while still decoupling as desired.
Here we demonstrate this advantage by applying the Magnus implementation using the first-order Euler step-method.
